---
title: "DrakeParameterResults11012017"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

1. [MCMC analysis](#MCMC)
2. [Hemlock](#hem)
3. [Ash](#ash)
4. [Oak](#oak)
5. [ALL](#all)

```{r setup, echo=FALSE,message=FALSE,result='hide'}
require(shiny)
library(rjags)
library(R2WinBUGS)
library(rbugs)
library(drc)
library(knitr)

setwd("~/Dropbox/BU2017/CH3/parameterest")
drake=read.csv(file="drake2013_trimmed.csv")
drake=na.omit(drake)
drakeRx=drake[-drake$activity<0,]

AP=drake[drake$substrate=="AP ",]#802 rows
AP=AP[-AP$activity<0,]
BG=drake[drake$substrate=="BG ",] #810 rows
BG=BG[-BG$activity<0,]
CBH=drake[drake$substrate=="CBH",] #803 rows
CBH=CBH[-CBH$activity<0,]
NAG=drake[drake$substrate=="NAG",] #810 rows 
NAG=NAG[-NAG$activity<0,]

APHEM=drake[drake$substrate=="AP "& drake$species=="TSCA",]
APHEM=APHEM[-APHEM$activity<0,]
APASH=drake[drake$substrate=="AP "& drake$species=="FRAM",]
APASH=APASH[-APASH$activity<0,]
APOAK=drake[drake$substrate=="AP "& drake$species=="QURU",]
APOAK=APOAK[-APOAK$activity<0,]

BGASH=drake[drake$substrate=="BG "& drake$species=="FRAM",]
BGASH=BGASH[-BGASH$activity<0,]
BGHEM=drake[drake$substrate=="BG "& drake$species=="TSCA",]
BGHEM=BGHEM[-BGHEM$activity<0,]
BGOAK=drake[drake$substrate=="BG "& drake$species=="QURU",]
BGOAK=BGOAK[-BGOAK$activity<0,]

CBHHEM=drake[drake$substrate=="CBH"& drake$species=="TSCA",]
CBHHEM=CBHHEM[-CBHHEM$activity<0,]
CBHASH=drake[drake$substrate=="CBH"& drake$species=="FRAM",]
CBHASH=CBHASH[-CBHASH$activity<0,]
CBHOAK=drake[drake$substrate=="CBH"& drake$species=="QURU",]
CBHOAK=CBHOAK[-CBHOAK$activity<0,]

NAGHEM=drake[drake$substrate=="NAG"& drake$species=="TSCA",]
NAGHEM=NAGHEM[-NAGHEM$activity<0,]
NAGASH=drake[drake$substrate=="NAG"& drake$species=="FRAM",]
NAGASH=NAGASH[-NAGASH$activity<0,]
NAGOAK=drake[drake$substrate=="NAG"& drake$species=="QURU",]
NAGOAK=NAGOAK[-NAGOAK$activity<0,]
#Vm for AP expected to be close to 3 and kM close to 100
#hist(rgamma(n=100000,shape=0.09,rate=0.10))
subs=c(0:1000)
```

#MCMC<a id="MCMC"></a>
MCMC was run in JAGS for each stand type and enzyme separately using the following format:

A michaelis menten process model for the reaction rate (Rx) specified by Vm, Km and substrate concentration (Sx).

> ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) 

A gamma distribution for the data model, speficied by the shape parameter and the rate parameter. The rate parameter of the data model is equal to the shape over the mean. This could also be done by having a separate specification for rate itself as a second line.

> Rx[i]~dgamma(shape,shape/ERx[i]) #data model

Priors on Vm, Km and the shape parameter using the gamma distribution.

> Vm ~ dgamma(2,0.5)

> Km ~ dgamma(6,0.04)

> shape ~ dgamma(0.1,0.1)

Three chains were used for each MCMC, with the following initial conditions: (Km=85,Vm=0.2), (Km=92,Vm=0.3), (Km=90,Vm=0.4). We used 10000 iterations thinned by 10.

Additionally, an MLE curve-fitting package was used to compare model fits.

##HEMLOCK <a id="hem"></a>

###AP HEM

```{r, echo=FALSE,cache=TRUE}
MMAPHEM = "
model {
for(i in 1:262){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = APHEM$activity
Sx = APHEM$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelAPHEM   <- jags.model (file = textConnection(MMAPHEM),
                         data = data,
                         inits = init.cond, n.chains = 3)
update(object = j.modelAPHEM,n.iter = 10000)
MMAPHEM   <- coda.samples (model = j.modelAPHEM,
                      variable.names = c("Vm","Km"),
                      n.iter = 10000, thin=10)
APHEMkm=summary(MMAPHEM)$statistics[1,1]
APHEMvm=summary(MMAPHEM)$statistics[2,1]

APHEMdrfit=drm(APHEM$activity~APHEM$concentration,fct=MM.2())
APHEMVMfit=summary(APHEMdrfit)$coefficients[1,1] 
APHEMKMfit=summary(APHEMdrfit)$coefficients[2,1]

plot(APHEM$concentration,APHEM$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="APHEM")
lines(subs,subs*APHEMvm/(APHEMkm+subs),lty=1,col="red")
lines(subs,subs*APHEMVMfit/(APHEMKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(APHEM$concentration),median(APHEM$activity[APHEM$concentration==max(APHEM$concentration)]),pch=16, col="blue")
points(max(APHEM$concentration),mean(APHEM$activity[APHEM$concentration==max(APHEM$concentration)]),pch=16, col="red")
#summary(APHEM)
```

###BG HEM

```{r, echo=FALSE,cache=TRUE}
MMBGHEM = "
model {
for(i in 1:268){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = BGHEM$activity
Sx = BGHEM$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelBGHEM   <- jags.model (file = textConnection(MMBGHEM),
                              data = data,
                              inits = init.cond, n.chains = 3)
update(object = j.modelBGHEM,n.iter = 10000)
MMBGHEM   <- coda.samples (model = j.modelBGHEM,
                           variable.names = c("Vm","Km"),
                           n.iter = 10000, thin=10)
BGHEMkm=summary(MMBGHEM)$statistics[1,1]
BGHEMvm=summary(MMBGHEM)$statistics[2,1]

BGHEMdrfit=drm(BGHEM$activity~BGHEM$concentration,fct=MM.2())
BGHEMVMfit=summary(BGHEMdrfit)$coefficients[1,1] 
BGHEMKMfit=summary(BGHEMdrfit)$coefficients[2,1]

plot(BGHEM$concentration,BGHEM$activity, xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="BGHEM")
lines(subs,subs*BGHEMvm/(BGHEMkm+subs),lty=1,col="red")
lines(subs,subs*BGHEMVMfit/(BGHEMKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(BGHEM$concentration),median(BGHEM$activity[BGHEM$concentration==max(BGHEM$concentration)]),pch=16, col="blue")
points(max(BGHEM$concentration),mean(BGHEM$activity[BGHEM$concentration==max(BGHEM$concentration)]),pch=16, col="red")
#summary(BGHEM)
```

###CBH HEM

```{r,echo=FALSE,cache=TRUE}
MMCBHHEM = "
model {
for(i in 1:242){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = CBHHEM$activity
Sx = CBHHEM$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelCBHHEM   <- jags.model (file = textConnection(MMCBHHEM),
                              data = data,
                              inits = init.cond, n.chains = 3)
update(object = j.modelCBHHEM,n.iter = 10000)
MMCBHHEM   <- coda.samples (model = j.modelCBHHEM,
                           variable.names = c("Vm","Km"),
                           n.iter = 10000, thin=10)
CBHHEMkm=summary(MMCBHHEM)$statistics[1,1]
CBHHEMvm=summary(MMCBHHEM)$statistics[2,1]

CBHHEMdrfit=drm(CBHHEM$activity~CBHHEM$concentration,fct=MM.2())
CBHHEMVMfit=summary(CBHHEMdrfit)$coefficients[1,1] 
CBHHEMKMfit=summary(CBHHEMdrfit)$coefficients[2,1]

plot(CBHHEM$concentration,CBHHEM$activity, xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="CBHHEM")
lines(subs,subs*CBHHEMvm/(CBHHEMkm+subs),lty=1,col="red")
lines(subs,subs*CBHHEMVMfit/(CBHHEMKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(CBHHEM$concentration),median(CBHHEM$activity[CBHHEM$concentration==max(CBHHEM$concentration)]),pch=16, col="blue")
points(max(CBHHEM$concentration),mean(CBHHEM$activity[CBHHEM$concentration==max(CBHHEM$concentration)]),pch=16, col="red")
#summary(CBHHEM)
```

###NAG HEM

```{r,echo=FALSE,cache=TRUE}
MMNAGHEM = "
model {
for(i in 1:255){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = NAGHEM$activity
Sx = NAGHEM$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelNAGHEM   <- jags.model (file = textConnection(MMNAGHEM),
                              data = data,
                              inits = init.cond, n.chains = 3)
update(object = j.modelNAGHEM,n.iter = 10000)
MMNAGHEM   <- coda.samples (model = j.modelNAGHEM,
                           variable.names = c("Vm","Km"),
                           n.iter = 10000, thin=10)
NAGHEMkm=summary(MMNAGHEM)$statistics[1,1]
NAGHEMvm=summary(MMNAGHEM)$statistics[2,1]

NAGHEMdrfit=drm(NAGHEM$activity~NAGHEM$concentration,fct=MM.2())
NAGHEMVMfit=summary(NAGHEMdrfit)$coefficients[1,1] 
NAGHEMKMfit=summary(NAGHEMdrfit)$coefficients[2,1]

plot(NAGHEM$concentration,NAGHEM$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="NAGHEM")
lines(subs,subs*NAGHEMvm/(NAGHEMkm+subs),lty=1,col="red")
lines(subs,subs*NAGHEMVMfit/(NAGHEMKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(NAGHEM$concentration),median(NAGHEM$activity[NAGHEM$concentration==max(NAGHEM$concentration)]),pch=16, col="blue")
points(max(NAGHEM$concentration),mean(NAGHEM$activity[NAGHEM$concentration==max(NAGHEM$concentration)]),pch=16, col="red")
#summary(NAGHEM)
```

##ASH <a id="ash"></a>

###AP ASH

```{r, echo=FALSE,cache=TRUE}
MMAPASH = "
model {
for(i in 1:270){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = APASH$activity
Sx = APASH$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelAPASH   <- jags.model (file = textConnection(MMAPASH),
                              data = data,
                              inits = init.cond, n.chains = 3)
update(object = j.modelAPASH,n.iter = 10000)
MMAPASH   <- coda.samples (model = j.modelAPASH,
                           variable.names = c("Vm","Km"),
                           n.iter = 10000, thin=10)
APASHkm=summary(MMAPASH)$statistics[1,1]
APASHvm=summary(MMAPASH)$statistics[2,1]

APASHdrfit=drm(APASH$activity~APASH$concentration,fct=MM.2())
APASHVMfit=summary(APASHdrfit)$coefficients[1,1] 
APASHKMfit=summary(APASHdrfit)$coefficients[2,1]

plot(APASH$concentration,APASH$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="APASH")
lines(subs,subs*APASHvm/(APASHkm+subs),lty=1,col="red")
lines(subs,subs*APASHVMfit/(APASHKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(APASH$concentration),median(APASH$activity[APASH$concentration==max(APASH$concentration)]),pch=16, col="blue")
points(max(APASH$concentration),mean(APASH$activity[APASH$concentration==max(APASH$concentration)]),pch=16, col="red")
#summary(APASH)
```

###BG ASH

```{r, echo=FALSE,cache=TRUE}
MMBGASH = "
model {
for(i in 1:270){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = BGASH$activity
Sx = BGASH$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelBGASH   <- jags.model (file = textConnection(MMBGASH),
                              data = data,
                              inits = init.cond, n.chains = 3)
update(object = j.modelBGASH,n.iter = 10000)
MMBGASH   <- coda.samples (model = j.modelBGASH,
                           variable.names = c("Vm","Km"),
                           n.iter = 10000, thin=10)
BGASHkm=summary(MMBGASH)$statistics[1,1]
BGASHvm=summary(MMBGASH)$statistics[2,1]

BGASHdrfit=drm(BGASH$activity~BGASH$concentration,fct=MM.2())
BGASHVMfit=summary(BGASHdrfit)$coefficients[1,1] 
BGASHKMfit=summary(BGASHdrfit)$coefficients[2,1]

plot(BGASH$concentration,BGASH$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="BGASH")
lines(subs,subs*BGASHvm/(BGASHkm+subs),lty=1,col="red")
lines(subs,subs*BGASHVMfit/(BGASHKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(BGASH$concentration),median(BGASH$activity[BGASH$concentration==max(BGASH$concentration)]),pch=16, col="blue")
points(max(BGASH$concentration),mean(BGASH$activity[BGASH$concentration==max(BGASH$concentration)]),pch=16, col="red")
#summary(BGASH)
```

###CBH ASH

```{r,echo=FALSE,cache=TRUE}
MMCBHASH = "
model {
for(i in 1:254){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = CBHASH$activity
Sx = CBHASH$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelCBHASH   <- jags.model (file = textConnection(MMCBHASH),
                               data = data,
                               inits = init.cond, n.chains = 3)
update(object = j.modelCBHASH,n.iter = 10000)
MMCBHASH   <- coda.samples (model = j.modelCBHASH,
                            variable.names = c("Vm","Km"),
                            n.iter = 10000, thin=10)
CBHASHkm=summary(MMCBHASH)$statistics[1,1]
CBHASHvm=summary(MMCBHASH)$statistics[2,1]

CBHASHdrfit=drm(CBHASH$activity~CBHASH$concentration,fct=MM.2())
CBHASHVMfit=summary(CBHASHdrfit)$coefficients[1,1] 
CBHASHKMfit=summary(CBHASHdrfit)$coefficients[2,1]

plot(CBHASH$concentration,CBHASH$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="CBHASH")
lines(subs,subs*CBHASHvm/(CBHASHkm+subs),lty=1,col="red")
lines(subs,subs*CBHASHVMfit/(CBHASHKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(CBHASH$concentration),median(CBHASH$activity[CBHASH$concentration==max(CBHASH$concentration)]),pch=16, col="blue")
points(max(CBHASH$concentration),mean(CBHASH$activity[CBHASH$concentration==max(CBHASH$concentration)]),pch=16, col="red")
#summary(CBHASH)
```

###NAG ASH

```{r,echo=FALSE,cache=TRUE}
MMNAGASH = "
model {
for(i in 1:263){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = NAGASH$activity
Sx = NAGASH$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelNAGASH   <- jags.model (file = textConnection(MMNAGASH),
                               data = data,
                               inits = init.cond, n.chains = 3)
update(object = j.modelNAGASH,n.iter = 10000)
MMNAGASH   <- coda.samples (model = j.modelNAGASH,
                            variable.names = c("Vm","Km"),
                            n.iter = 10000, thin=10)
NAGASHkm=summary(MMNAGASH)$statistics[1,1]
NAGASHvm=summary(MMNAGASH)$statistics[2,1]

NAGASHdrfit=drm(NAGASH$activity~NAGASH$concentration,fct=MM.2())
NAGASHVMfit=summary(NAGASHdrfit)$coefficients[1,1] 
NAGASHKMfit=summary(NAGASHdrfit)$coefficients[2,1]

plot(NAGASH$concentration,NAGASH$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="NAGASH")
lines(subs,subs*NAGASHvm/(NAGASHkm+subs),lty=1,col="red")
lines(subs,subs*NAGASHVMfit/(NAGASHKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(NAGASH$concentration),median(NAGASH$activity[NAGASH$concentration==max(NAGASH$concentration)]),pch=16, col="blue")
points(max(NAGASH$concentration),mean(NAGASH$activity[NAGASH$concentration==max(NAGASH$concentration)]),pch=16, col="red")
#summary(NAGASH)
```

##OAK <a id="oak"></a>

###AP OAK

```{r, echo=FALSE,cache=TRUE}
MMAPOAK = "
model {
for(i in 1:270){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = APOAK$activity
Sx = APOAK$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelAPOAK   <- jags.model (file = textConnection(MMAPOAK),
                              data = data,
                              inits = init.cond, n.chains = 3)
update(object = j.modelAPOAK,n.iter = 10000)
MMAPOAK   <- coda.samples (model = j.modelAPOAK,
                           variable.names = c("Vm","Km"),
                           n.iter = 10000, thin=10)
APOAKkm=summary(MMAPOAK)$statistics[1,1]
APOAKvm=summary(MMAPOAK)$statistics[2,1]

APOAKdrfit=drm(APOAK$activity~APOAK$concentration,fct=MM.2())
APOAKVMfit=summary(APOAKdrfit)$coefficients[1,1] 
APOAKKMfit=summary(APOAKdrfit)$coefficients[2,1]

plot(APOAK$concentration,APOAK$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="APOAK")
lines(subs,subs*APOAKvm/(APOAKkm+subs),lty=1,col="red")
lines(subs,subs*APOAKVMfit/(APOAKKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(APOAK$concentration),median(APOAK$activity[APOAK$concentration==max(APOAK$concentration)]),pch=16, col="blue")
points(max(APOAK$concentration),mean(APOAK$activity[APOAK$concentration==max(APOAK$concentration)]),pch=16, col="red")
#summary(APOAK)
```

###BG OAK

```{r, echo=FALSE,cache=TRUE}
MMBGOAK = "
model {
for(i in 1:270){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = BGOAK$activity
Sx = BGOAK$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelBGOAK   <- jags.model (file = textConnection(MMBGOAK),
                              data = data,
                              inits = init.cond, n.chains = 3)
update(object = j.modelBGOAK,n.iter = 10000)
MMBGOAK   <- coda.samples (model = j.modelBGOAK,
                           variable.names = c("Vm","Km"),
                           n.iter = 10000, thin=10)
BGOAKkm=summary(MMBGOAK)$statistics[1,1]
BGOAKvm=summary(MMBGOAK)$statistics[2,1]

BGOAKdrfit=drm(BGOAK$activity~BGOAK$concentration,fct=MM.2())
BGOAKVMfit=summary(BGOAKdrfit)$coefficients[1,1] 
BGOAKKMfit=summary(BGOAKdrfit)$coefficients[2,1]

plot(BGOAK$concentration,BGOAK$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="BGOAK")
lines(subs,subs*BGOAKvm/(BGOAKkm+subs),lty=1,col="red")
lines(subs,subs*BGOAKVMfit/(BGOAKKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(BGOAK$concentration),median(BGOAK$activity[BGOAK$concentration==max(BGOAK$concentration)]),pch=16, col="blue")
points(max(BGOAK$concentration),mean(BGOAK$activity[BGOAK$concentration==max(BGOAK$concentration)]),pch=16, col="red")
#summary(BGOAK)
```

###CBH OAK

```{r,echo=FALSE,cache=TRUE}
MMCBHOAK = "
model {
for(i in 1:257){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = CBHOAK$activity
Sx = CBHOAK$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelCBHOAK   <- jags.model (file = textConnection(MMCBHOAK),
                               data = data,
                               inits = init.cond, n.chains = 3)
update(object = j.modelCBHOAK,n.iter = 10000)
MMCBHOAK   <- coda.samples (model = j.modelCBHOAK,
                            variable.names = c("Vm","Km"),
                            n.iter = 10000, thin=10)
CBHOAKkm=summary(MMCBHOAK)$statistics[1,1]
CBHOAKvm=summary(MMCBHOAK)$statistics[2,1]

#CBHOAKdrfit=drm(CBHOAK$activity~CBHOAK$concentration,fct=MM.2())
#CBHOAKVMfit=summary(CBHOAKdrfit)$coefficients[1,1] 
#CBHOAKKMfit=summary(CBHOAKdrfit)$coefficients[2,1]

plot(CBHOAK$concentration,CBHOAK$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="CBHOAK")
lines(subs,subs*CBHOAKvm/(CBHOAKkm+subs),lty=1,col="red")
#lines(subs,subs*CBHOAKVMfit/(CBHOAKKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(CBHOAK$concentration),median(CBHOAK$activity[CBHOAK$concentration==max(CBHOAK$concentration)]),pch=16, col="blue")
points(max(CBHOAK$concentration),mean(CBHOAK$activity[CBHOAK$concentration==max(CBHOAK$concentration)]),pch=16, col="red")
#summary(CBHOAK)
```

###NAG OAK

```{r,echo=FALSE,cache=TRUE}
MMNAGOAK = "
model {
for(i in 1:264){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = NAGOAK$activity
Sx = NAGOAK$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelNAGOAK   <- jags.model (file = textConnection(MMNAGOAK),
                               data = data,
                               inits = init.cond, n.chains = 3)
update(object = j.modelNAGOAK,n.iter = 10000)
MMNAGOAK   <- coda.samples (model = j.modelNAGOAK,
                            variable.names = c("Vm","Km"),
                            n.iter = 10000, thin=10)
NAGOAKkm=summary(MMNAGOAK)$statistics[1,1]
NAGOAKvm=summary(MMNAGOAK)$statistics[2,1]

NAGOAKdrfit=drm(NAGOAK$activity~NAGOAK$concentration,fct=MM.2())
NAGOAKVMfit=summary(NAGOAKdrfit)$coefficients[1,1] 
NAGOAKKMfit=summary(NAGOAKdrfit)$coefficients[2,1]

plot(NAGOAK$concentration,NAGOAK$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="NAGOAK")
lines(subs,subs*NAGOAKvm/(NAGOAKkm+subs),lty=1,col="red")
lines(subs,subs*NAGOAKVMfit/(NAGOAKKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(NAGOAK$concentration),median(NAGOAK$activity[NAGOAK$concentration==max(NAGOAK$concentration)]),pch=16, col="blue")
points(max(NAGOAK$concentration),mean(NAGOAK$activity[NAGOAK$concentration==max(NAGOAK$concentration)]),pch=16, col="red")
#summary(NAGOAK)
```

#ENZYMES ACROSS STANDS <a id="all"></a>

###NAG

```{r,echo=FALSE,cache=TRUE}
MMNAG = "
model {
for(i in 1:782){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = NAG$activity
Sx = NAG$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelNAG   <- jags.model (file = textConnection(MMNAG),
                               data = data,
                               inits = init.cond, n.chains = 3)
update(object = j.modelNAG,n.iter = 10000)
MMNAG   <- coda.samples (model = j.modelNAG,
                            variable.names = c("Vm","Km"),
                            n.iter = 10000, thin=10)
NAGkm=summary(MMNAG)$statistics[1,1]
NAGvm=summary(MMNAG)$statistics[2,1]

NAGdrfit=drm(NAGOAK$activity~NAGOAK$concentration,fct=MM.2())
NAGVMfit=summary(NAGdrfit)$coefficients[1,1] 
NAGKMfit=summary(NAGdrfit)$coefficients[2,1]

plot(NAG$concentration,NAG$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="NAG")
lines(subs,subs*NAGvm/(NAGkm+subs),lty=1,col="red")
lines(subs,subs*NAGVMfit/(NAGKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(NAG$concentration),median(NAG$activity[NAG$concentration==max(NAG$concentration)]),pch=16, col="blue")
points(max(NAG$concentration),mean(NAG$activity[NAG$concentration==max(NAG$concentration)]),pch=16, col="red")
#summary(NAG)
```

###CBH

```{r,echo=FALSE,cache=TRUE}
MMCBH = "
model {
for(i in 1:753){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = CBH$activity
Sx = CBH$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelCBH   <- jags.model (file = textConnection(MMCBH),
                            data = data,
                            inits = init.cond, n.chains = 3)
update(object = j.modelCBH,n.iter = 10000)
MMCBH   <- coda.samples (model = j.modelCBH,
                         variable.names = c("Vm","Km"),
                         n.iter = 10000, thin=10)
CBHkm=summary(MMCBH)$statistics[1,1]
CBHvm=summary(MMCBH)$statistics[2,1]

#CBHdrfit=drm(CBHOAK$activity~CBHOAK$concentration,fct=MM.2())
#CBHVMfit=summary(CBHdrfit)$coefficients[1,1] 
#CBHKMfit=summary(CBHdrfit)$coefficients[2,1]

plot(CBH$concentration,CBH$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="CBH")
lines(subs,subs*CBHvm/(CBHkm+subs),lty=1,col="red")
#lines(subs,subs*CBHVMfit/(CBHKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(CBH$concentration),median(CBH$activity[CBH$concentration==max(CBH$concentration)]),pch=16, col="blue")
points(max(CBH$concentration),mean(CBH$activity[CBH$concentration==max(CBH$concentration)]),pch=16, col="red")
#summary(CBH)
```

###BG

```{r,echo=FALSE,cache=TRUE}
MMBG = "
model {
for(i in 1:808){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = BG$activity
Sx = BG$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelBG   <- jags.model (file = textConnection(MMBG),
                            data = data,
                            inits = init.cond, n.chains = 3)
update(object = j.modelBG,n.iter = 10000)
MMBG   <- coda.samples (model = j.modelBG,
                         variable.names = c("Vm","Km"),
                         n.iter = 10000, thin=10)
BGkm=summary(MMBG)$statistics[1,1]
BGvm=summary(MMBG)$statistics[2,1]

#BGdrfit=drm(BGOAK$activity~BGOAK$concentration,fct=MM.2())
#BGVMfit=summary(BGdrfit)$coefficients[1,1] 
#BGKMfit=summary(BGdrfit)$coefficients[2,1]

plot(BG$concentration,BG$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="BG")
lines(subs,subs*BGvm/(BGkm+subs),lty=1,col="red")
#lines(subs,subs*BGVMfit/(BGKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(BG$concentration),median(BG$activity[BG$concentration==max(BG$concentration)]),pch=16, col="blue")
points(max(BG$concentration),mean(BG$activity[BG$concentration==max(BG$concentration)]),pch=16, col="red")
#summary(BG)
```

###AP

```{r,echo=FALSE,cache=TRUE}
MMAP = "
model {
for(i in 1:802){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = AP$activity
Sx = AP$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelAP   <- jags.model (file = textConnection(MMAP),
                           data = data,
                           inits = init.cond, n.chains = 3)
update(object = j.modelAP,n.iter = 10000)
MMAP   <- coda.samples (model = j.modelAP,
                        variable.names = c("Vm","Km"),
                        n.iter = 10000, thin=10)
APkm=summary(MMAP)$statistics[1,1]
APvm=summary(MMAP)$statistics[2,1]

#APdrfit=drm(APOAK$activity~APOAK$concentration,fct=MM.2())
#APVMfit=summary(APdrfit)$coefficients[1,1] 
#APKMfit=summary(APdrfit)$coefficients[2,1]

plot(AP$concentration,AP$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="AP")
lines(subs,subs*APvm/(APkm+subs),lty=1,col="red")
#lines(subs,subs*APVMfit/(APKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(AP$concentration),median(AP$activity[AP$concentration==max(AP$concentration)]),pch=16, col="blue")
points(max(AP$concentration),mean(AP$activity[AP$concentration==max(AP$concentration)]),pch=16, col="red")
#summary(AP)
```

###SINGLE MODEL

```{r,echo=FALSE,cache=TRUE}
MMALL= "
model {
for(i in 1:3145){
ERx[i]<- Vm*Sx[i]/(Km+Sx[i]) #process model
Rx[i]~dgamma(shape,shape/ERx[i]) #data model
}
Vm ~ dgamma(2,0.5) 
Km ~ dgamma(6,0.04)
shape ~ dgamma(0.1,0.1)
}"

Rx = drakeRx$activity
Sx = drakeRx$concentration

data = list(Rx=Rx,Sx=Sx)
init.cond <- list()
init.cond[[1]] = list(Km=85,Vm=0.2)
init.cond[[2]] = list(Km=92,Vm=0.3)
init.cond[[3]] = list(Km=90,Vm=0.4)

#compile JAGS model <<<<<<<<< 
j.modelALL   <- jags.model (file = textConnection(MMALL),
                            data = data,
                            inits = init.cond, n.chains = 3)
update(object = j.modelALL,n.iter = 10000)
MMALL   <- coda.samples (model = j.modelALL,
                         variable.names = c("Vm","Km"),
                         n.iter = 10000, thin=10)
ALLkm=summary(MMALL)$statistics[1,1]
ALLvm=summary(MMALL)$statistics[2,1]

#ALLdrfit=drm(drakeRx$activity~drakeRx$concentration,fct=MM.2())
#ALLVMfit=summary(ALLdrfit)$coefficients[1,1] 
#ALLKMfit=summary(ALLdrfit)$coefficients[2,1]

plot(drakeRx$concentration,drakeRx$activity,xlab="Concentration (uM)",ylab="Activity (umol g-1 hr-1)",main="ALL")
lines(subs,subs*ALLvm/(ALLkm+subs),lty=1,col="red")
#lines(subs,subs*ALLVMfit/(ALLKMfit+subs),lty=2,col="black")
legend("topleft",ncol=2,legend=c("MLE","MCMC","Median","Mean"),col=c("black","red","blue","red"),lty=c(2,1,NA,NA),pch=c(NA,NA,16,16),bty="n")
points(max(drakeRx$concentration),median(drakeRx$activity[drakeRx$concentration==max(drakeRx$concentration)]),pch=16, col="blue")
points(max(drakeRx$concentration),mean(drakeRx$activity[drakeRx$concentration==max(drakeRx$concentration)]),pch=16, col="red")
#summary(MMALL)
```

#SUMMARY TABLES

**CBH**

```{r,echo=FALSE}
CBHsummary=matrix(NA,3,4)
colnames(CBHsummary)=c("KM-MLE","KM-MCMC","VM-MLE","VM-MCMC")
rownames(CBHsummary)=c("HEM","ASH","OAK")
CBHsummary[1,1]=CBHHEMKMfit
CBHsummary[1,2]=CBHHEMkm
CBHsummary[1,3]=CBHHEMVMfit
CBHsummary[1,4]=CBHHEMvm
CBHsummary[2,1]=CBHASHKMfit
CBHsummary[2,2]=CBHASHkm
CBHsummary[2,3]=CBHASHVMfit
CBHsummary[2,4]=CBHASHvm

CBHsummary[3,2]=CBHOAKkm

CBHsummary[3,4]=CBHOAKvm
kable(CBHsummary)
```

**NAG**

```{r,echo=FALSE}
NAGsummary=matrix(NA,3,4)
colnames(NAGsummary)=c("KM-MLE","KM-MCMC","VM-MLE","VM-MCMC")
rownames(NAGsummary)=c("HEM","ASH","OAK")
NAGsummary[1,1]=NAGHEMKMfit
NAGsummary[1,2]=NAGHEMkm
NAGsummary[1,3]=NAGHEMVMfit
NAGsummary[1,4]=NAGHEMvm
NAGsummary[2,1]=NAGASHKMfit
NAGsummary[2,2]=NAGASHkm
NAGsummary[2,3]=NAGASHVMfit
NAGsummary[2,4]=NAGASHvm
NAGsummary[3,1]=NAGOAKKMfit
NAGsummary[3,2]=NAGOAKkm
NAGsummary[3,3]=NAGOAKVMfit
NAGsummary[3,4]=NAGOAKvm
kable(NAGsummary)
```

**BG**

```{r,echo=FALSE}
BGsummary=matrix(NA,3,4)
colnames(BGsummary)=c("KM-MLE","KM-MCMC","VM-MLE","VM-MCMC")
rownames(BGsummary)=c("HEM","ASH","OAK")
BGsummary[1,1]=BGHEMKMfit
BGsummary[1,2]=BGHEMkm
BGsummary[1,3]=BGHEMVMfit
BGsummary[1,4]=BGHEMvm
BGsummary[2,1]=BGASHKMfit
BGsummary[2,2]=BGASHkm
BGsummary[2,3]=BGASHVMfit
BGsummary[2,4]=BGASHvm
BGsummary[3,1]=BGOAKKMfit
BGsummary[3,2]=BGOAKkm
BGsummary[3,3]=BGOAKVMfit
BGsummary[3,4]=BGOAKvm
kable(BGsummary)
```

**AP**

```{r,echo=FALSE}
APsummary=matrix(NA,3,4)
colnames(APsummary)=c("KM-MLE","KM-MCMC","VM-MLE","VM-MCMC")
rownames(APsummary)=c("HEM","ASH","OAK")
APsummary[1,1]=APHEMKMfit
APsummary[1,2]=APHEMkm
APsummary[1,3]=APHEMVMfit
APsummary[1,4]=APHEMvm
APsummary[2,1]=APASHKMfit
APsummary[2,2]=APASHkm
APsummary[2,3]=APASHVMfit
APsummary[2,4]=APASHvm
APsummary[3,1]=APOAKKMfit
APsummary[3,2]=APOAKkm
APsummary[3,3]=APOAKVMfit
APsummary[3,4]=APOAKvm
kable(APsummary)
```

```{r,echo=FALSE}
combosummary=matrix(NA,5,2)
colnames(combosummary)=c("KM","VM")
rownames(combosummary)=c("CBH","NAG","AP","BG","ALL")
combosummary[1,1]=CBHkm
combosummary[1,2]=CBHvm
combosummary[2,1]=NAGkm
combosummary[2,2]=NAGvm
combosummary[3,1]=APkm
combosummary[3,2]=APvm
combosummary[4,1]=BGkm
combosummary[4,2]=BGvm
combosummary[5,1]=ALLkm
combosummary[5,2]=ALLvm
kable(combosummary)
```